---
title: "Body mass of Palmer penguins by species and sex"
subtitle: "A worked example of the R analysis project template"
date: today
date-format: long
format:
  # Editable copy of slides.qmd: the same content with tables in place of
  # the cards, on the template slides-reference.pptx
  pptx:
    reference-doc: slides-reference.pptx
    output-file: slides.pptx
---

```{r}
#| label: setup
suppressPackageStartupMessages({
  library(here)
  library(data.table)
})
here::i_am("reports-src/slides-pptx.qmd")
source(here("R/utils/config.R"))
load_config()
invisible(list2env(
  readRDS(get_data_path("outputs", "summary_env")), environment()
))

# Formatting only: every number comes from summary_env.rds (03)
num.fn <- function(x) sub("^-", "\u2212", sprintf("%.0f", x))
ci.fn <- function(x) {
  sprintf("%s [%s, %s]", num.fn(x$estimate), num.fn(x$lower),
          num.fn(x$upper))
}
difference.fn <- function(s) differences[species == s]
n.fn <- function(s) means[species == s, sum(n)]
spread <- paste(num.fn(min(differences$estimate)), "to",
                num.fn(max(differences$estimate)), "g")
dropped.v <- data$dropped_by_species[, sprintf("%d %s", N, species)]

# Claims the slides make, checked against the numbers
stopifnot(
  all(differences$lower > 0),
  model$chinstrap$lower < 0, model$chinstrap$upper > 0
)
```

## Question and data

| Question | How much heavier are male penguins than females, and does the difference depend on the species? |
|-----|---------------|
| **Penguins measured** | `r data$penguins` penguins of three species, Palmer Archipelago, Antarctica, `r data$years[1]` to `r data$years[2]` |
| **Analysed** | `r data$complete` with every measurement and a recorded sex; left out: `r paste(dropped.v, collapse = " and ")` |
| **Data** | `datasets::penguins`, shipped with R since 4.5; Gorman, Williams and Fraser (2014) |

## Males are heavier in every species

Points: individual penguins. Black: means with `r 100 * ci_level`%
confidence intervals.

```{r}
knitr::include_graphics(slide_figure)
```

## Male minus female body mass

Difference in means within each species, with its Welch confidence
interval at `r 100 * ci_level`%.

| | Adelie | Chinstrap | Gentoo |
|---|---|---|---|
| **Difference (g)** | `r ci.fn(difference.fn("Adelie"))` | `r ci.fn(difference.fn("Chinstrap"))` | `r ci.fn(difference.fn("Gentoo"))` |
| **Penguins** | `r n.fn("Adelie")` | `r n.fn("Chinstrap")` | `r n.fn("Gentoo")` |

## Adjusted for species

- **Males** heavier by `r ci.fn(model$male)` g
- **Gentoo** heavier than Adelie of the same sex,
  by `r ci.fn(model$gentoo)` g
- **Chinstrap** and Adelie: `r ci.fn(model$chinstrap)` g, no clear
  difference
- **Model:** body mass on species and sex, `r model$n` penguins,
  R² `r sprintf("%.2f", model$r_squared)`

## Caveats

- **One difference for all species:** the model assumes it; the
  within-species differences range from `r spread`
- **Complete records only:** `r data$dropped` penguins without a recorded
  sex are left out
