---
title: "Body mass of Palmer penguins by species and sex"
subtitle: "A worked example of the R analysis project template"
date: today
date-format: long
format:
  # Editable copy of slides.qmd: the same content with tables in place of
  # the cards, on the template slides-reference.pptx
  pptx:
    reference-doc: slides-reference.pptx
    output-file: slides.pptx
---

```{r}
#| label: setup
suppressPackageStartupMessages({
  library(here)
  library(data.table)
})
here::i_am("reports-src/slides-pptx.qmd")
source(here("R/utils/config.R"))
load_config()
invisible(list2env(
  readRDS(get_data_path("outputs", "summary_env")), environment()
))

# Formatting only: every number comes from summary_env.rds (03)
num.fn <- function(x) sub("^-", "\u2212", sprintf("%.0f", x))
range.fn <- function(x) {
  paste(num.fn(x$lower), "to", num.fn(x$upper))
}
difference.fn <- function(s) differences[species == s]
n.fn <- function(s) means[species == s, sum(n)]
largest <- differences[which.max(estimate), species]
spread <- paste(num.fn(min(differences$estimate)), "to",
                num.fn(max(differences$estimate)), "g")
dropped.v <- data$dropped_by_species[, sprintf("%d %s", N, species)]
ci_pct <- 100 * ci_level

# Claims the slides make, checked against the numbers
stopifnot(
  all(differences$lower > 0),
  model$male$lower > 0,
  model$chinstrap$lower < 0, model$chinstrap$upper > 0
)
```

## How much heavier are males, and does it depend on the species?

| | |
|-----|---------------|
| **Penguins measured** | `r data$penguins` penguins of three species, Palmer Archipelago, Antarctica, `r data$years[1]` to `r data$years[2]` |
| **Analysed** | `r data$complete` with every measurement and a recorded sex; left out: `r paste(dropped.v, collapse = " and ")` |
| **Data** | `datasets::penguins`, shipped with R since 4.5; Gorman, Williams and Fraser (2014) |

## Males are heavier in every species

`r data$complete` penguins. Points: individual penguins. Black: means with
`r ci_pct`% confidence intervals.

```{r}
knitr::include_graphics(slide_figure)
```

## The difference is largest in `r largest` penguins

Male minus female mean body mass within each species, with a Welch
confidence interval (CI).

| | Adelie | Chinstrap | Gentoo |
|---|---|---|---|
| **Difference (g)** | `r num.fn(difference.fn("Adelie")$estimate)` | `r num.fn(difference.fn("Chinstrap")$estimate)` | `r num.fn(difference.fn("Gentoo")$estimate)` |
| **`r ci_pct`% CI (g)** | `r range.fn(difference.fn("Adelie"))` | `r range.fn(difference.fn("Chinstrap"))` | `r range.fn(difference.fn("Gentoo"))` |
| **Penguins** | `r n.fn("Adelie")` | `r n.fn("Chinstrap")` | `r n.fn("Gentoo")` |

## Adjusted for species, males are heavier by `r num.fn(model$male$estimate)` g

Linear model of body mass on species and sex; `r model$n` penguins;
R² `r sprintf("%.2f", model$r_squared)`. Chinstrap and Adelie penguins of
the same sex do not clearly differ. The model assumes one difference for
all species, while within species it ranges from `r spread`.

| Comparison | Difference (g) | `r ci_pct`% CI (g) |
|---|---|---|
| Males vs females | `r num.fn(model$male$estimate)` | `r range.fn(model$male)` |
| Chinstrap vs Adelie | `r num.fn(model$chinstrap$estimate)` | `r range.fn(model$chinstrap)` |
| Gentoo vs Adelie | `r num.fn(model$gentoo$estimate)` | `r range.fn(model$gentoo)` |
