---
title: "Body mass of Palmer penguins by species and sex"
subtitle: "A worked example of the R analysis project template"
date: today
date-format: long
format:
  revealjs:
    embed-resources: true
    theme: [default, slides.scss]
    slide-number: c/t
    menu: false
    width: 1600
    height: 900
    margin: 0.06
---

```{r}
#| label: setup
suppressPackageStartupMessages({
  library(here)
  library(data.table)
})
here::i_am("reports-src/slides.qmd")
source(here("R/utils/config.R"))
load_config()
invisible(list2env(
  readRDS(get_data_path("outputs", "summary_env")), environment()
))

# Formatting only: every number comes from summary_env.rds (03)
num.fn <- function(x) sub("^-", "\u2212", sprintf("%.0f", x))
ci.fn <- function(x) {
  sprintf("%s [%s, %s]", num.fn(x$estimate), num.fn(x$lower),
          num.fn(x$upper))
}
difference.fn <- function(s) differences[species == s]
est.fn <- function(s) paste(num.fn(difference.fn(s)$estimate), "g")
range.fn <- function(s) {
  paste(num.fn(difference.fn(s)$lower), "to", num.fn(difference.fn(s)$upper),
        "g")
}
n.fn <- function(s) means[species == s, sum(n)]
spread <- paste(num.fn(min(differences$estimate)), "to",
                num.fn(max(differences$estimate)), "g")
dropped.v <- data$dropped_by_species[, sprintf("%d %s", N, species)]

# Claims the slides make, checked against the numbers
stopifnot(
  all(differences$lower > 0),
  model$chinstrap$lower < 0, model$chinstrap$upper > 0
)
```

## Question and data

[How much heavier are male penguins than females, and does the difference
depend on the species?]{.lead}

::: {.cards}
::: {.card .red}
**Penguins measured**

[`r data$penguins`]{.big} [penguins, three species]{.unit}

Palmer Archipelago, Antarctica, `r data$years[1]` to `r data$years[2]`
:::

::: {.card .red}
**Analysed**

[`r data$complete`]{.big} [with every measurement and a recorded sex]{.unit}

Left out: `r paste(dropped.v, collapse = " and ")`
:::
:::

::: {.note}
**Data:** `datasets::penguins`, shipped with R since 4.5; Gorman, Williams
and Fraser (2014).
:::

## Males are heavier in every species

```{r}
#| fig-align: center
#| fig-cap: !expr 'sprintf("Points: individual penguins. Black: means with %.0f%% confidence intervals.", 100 * ci_level)'
knitr::include_graphics(slide_figure)
```

## Male minus female body mass

::: {.cards}
::: {.card .adelie}
**Adelie**

[`r est.fn("Adelie")`]{.big} [`r n.fn("Adelie")` penguins]{.unit}

CI `r range.fn("Adelie")`
:::

::: {.card .chinstrap}
**Chinstrap**

[`r est.fn("Chinstrap")`]{.big} [`r n.fn("Chinstrap")` penguins]{.unit}

CI `r range.fn("Chinstrap")`
:::

::: {.card .gentoo}
**Gentoo**

[`r est.fn("Gentoo")`]{.big} [`r n.fn("Gentoo")` penguins]{.unit}

CI `r range.fn("Gentoo")`
:::
:::

::: {.note}
**Within each species:** difference in means, with a Welch confidence
interval (CI) of `r 100 * ci_level`%.
:::

## Adjusted for species

- **Males** heavier by `r ci.fn(model$male)` g
- **Gentoo** heavier than Adelie of the same sex,
  by `r ci.fn(model$gentoo)` g
- **Chinstrap** and Adelie: `r ci.fn(model$chinstrap)` g, no clear
  difference
- **Model:** body mass on species and sex, `r model$n` penguins,
  R² `r sprintf("%.2f", model$r_squared)`

## Caveats

- **One difference for all species:** the model assumes it; the
  within-species differences range from `r spread`
- **Complete records only:** `r data$dropped` penguins without a recorded
  sex are left out
