---
title: "Body mass of Palmer penguins by species and sex"
subtitle: "A worked example of the R analysis project template"
date: today
bibliography: references.bib
---

```{r}
#| label: setup
suppressPackageStartupMessages({
  library(here)
  library(data.table)
})
here::i_am("reports-src/summary.qmd")
source(here("R/utils/config.R"))
source(here("R/utils/tables.R"))
load_config()
invisible(list2env(
  readRDS(get_data_path("outputs", "summary_env")), environment()
))

# Formatting only: every number comes from summary_env.rds (03)
# A true minus sign, and a no-break space so a number keeps its unit
num.fn <- function(x) sub("^-", "\u2212", sprintf("%.0f", x))
ci.fn <- function(x) {
  sprintf("%s\u00a0g (%.0f%% CI %s to %s)", num.fn(x$estimate),
          100 * ci_level, num.fn(x$lower), num.fn(x$upper))
}
dropped.v <- data$dropped_by_species[, sprintf("%d %s", N, species)]
smallest.dt <- differences[which.min(estimate)]
largest.dt <- differences[which.max(estimate)]
r_squared_pct <- sprintf("%.0f%%", 100 * model$r_squared)

# Claims the text makes, checked against the numbers
stopifnot(
  all(differences$lower > 0),
  model$chinstrap$lower < 0, model$chinstrap$upper > 0,
  model$gentoo$lower > 0
)
```

# Purpose

This report shows how the parts of the project template fit together on
a small public dataset. Three scripts prepare the data, compare body mass
between species and sexes, and gather every number this report and the
slides use into one file. The report does no analysis of its own: every
number below is read from that file.

# Data

The data are measurements of adult penguins of three species near Palmer
Station, Antarctica, collected from `r data$years[1]` to `r data$years[2]`
[@gorman2014] and distributed with R's datasets package
[@horst2022; @kaye2025]. Of the `r data$penguins` penguins, a total
of `r data$complete` had every measurement and a recorded sex. The
other `r data$dropped` (`r paste(dropped.v, collapse = " and ")`) were
left out.

# Methods

Mean body mass is given for each species and sex with a confidence
interval of `r 100 * ci_level`% based on the t distribution. The
difference between males and females within each species comes with a
Welch interval. One linear model, body mass on species and sex, estimates
the difference between the sexes adjusted for species. It assumes the same
difference in every species; the within-species intervals show how far
that holds.

# Results

Males were heavier than females in all three species (@fig-body-mass,
@tbl-body-mass).
The difference was smallest in `r smallest.dt$species` penguins,
at `r ci.fn(smallest.dt)`, and largest in `r largest.dt$species`
penguins, at `r ci.fn(largest.dt)`. Adjusted for species, males were
heavier by `r ci.fn(model$male)`, and the
model explained `r r_squared_pct` of the variance in body mass.
Gentoo penguins were heavier than Adelie penguins of the same sex
by `r ci.fn(model$gentoo)`; Chinstrap and Adelie penguins differed
by `r ci.fn(model$chinstrap)`, an interval that includes zero.

```{r}
#| label: fig-body-mass
#| out-width: "100%"
#| fig-pos: H
#| fig-cap: !expr sprintf("**Males are heavier in every species.** Body mass of each penguin, with the mean and %.0f%% confidence interval for each species and sex (%d penguins).", 100 * ci_level, data$complete)
knitr::include_graphics(figure)
```

```{r}
#| label: tbl-body-mass
#| tbl-pos: H
#| tbl-cap: "**Body mass by species and sex.** Penguins with every measurement and a recorded sex."
create_body_mass_table.fn(
  body_mass_table, sex_labels.v = get_parameter("sex_labels"),
  ci_level = ci_level, title = NULL
) |>
  style_pdf_table.fn()
```

Because the within-species differences range
from `r num.fn(smallest.dt$estimate)` to `r num.fn(largest.dt$estimate)` g,
the adjusted difference is an average over species rather than a value
that holds in each of them.

# References {.unnumbered}
